---
title: "2024 UC Salary Cohort Classification and Crosswalk Audit"
author: "Robert J. Dellinger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    df_print: paged
---

<!-- 1. Title -->
<!-- Rendered through the YAML title block above. -->

<!-- 2. Name -->
<!-- Rendered through the YAML author field above. -->

<!-- 3. Date -->
<!-- Rendered through the YAML date field above. -->

<!-- 4. Objective -->
<!-- 5. Source -->
<!-- 6. Source documentation and information -->
<!-- 7. Methods, including equations and conceptual definitions -->

This report is the repository's cohort-classification audit. It takes the intentionally overlapping 2024 candidate pool produced by the focused scrape and turns it into an analysis-ready person-level cohort file with an explicit crosswalk, a cohort-rule reference, a title-reference join audit, and QA outputs that show how the recoding decisions were made.

> Interpretive note: this report is a classification artifact, not a substantive findings report. Its tables document how rows move from the upstream candidate pool into final analytic cohorts, and its academic title reference is a partial join aid rather than a universal dictionary for all University of California staff titles.

## Classification overview

The objective is to turn the 2024 Transparent California candidate pool into an analysis-ready salary dataset with explicit cohort definitions, title-reference joins where available, documented priority rules, and reproducible QA outputs that make the recoding process auditable.

### Source basis

The cleaning step uses the 2024 candidate pool, the standardized UC academic title reference, and the manual query registry that defines analytic cohorts, query priorities, and pattern-based inclusion and exclusion rules. In other words, this stage moves the repository from an intentionally overlapping scrape corpus to an explicit, auditable cohort-classification output.

### Source documentation and interpretation limits

The candidate pool contains intentional overlap because some search terms are broad and search pages can match names as well as job titles. The query registry therefore supplies the explicit cohort logic and priority ordering that resolve overlaps. The UC academic title reference covers academic titles but not all staff titles, so missing staff title codes are treated explicitly rather than guessed. In this repository, that reference file is a partial academic join aid rather than a universal dictionary for all University of California staff titles. The outputs of this stage should therefore be read as classification artifacts: they document how the repository moved from candidate rows to analytic cohorts and which titles matched the academic reference versus remaining unmatched.

### Classification logic

The workflow assigns cohorts in priority order and deduplicates by normalized employee name, normalized title, year, and agency. It keeps both source pay components and derived shares such as overtime share of total pay for later analysis. This document is the classification audit for the repository: it records the cleaned cohort file, the cohort reference table, the title crosswalk, the title-join audit, and the cleaning QA summary so downstream users can inspect how rows moved from overlapping candidate pools into final analytic cohorts. In methodological terms, this stage is where the repository locks the person-level cohort design that the cross-sectional report will later summarize.

<!-- 8. Setup and load packages -->

## Setup and load packages

This section loads helper functions, packages, and file paths.

```{r setup-classification-audit}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

utility_path <- if (file.exists("scripts/functions/uc_salary_utils.R")) {
  "scripts/functions/uc_salary_utils.R"
} else {
  "functions/uc_salary_utils.R"
}

source(utility_path)
uc_salary_load_packages()

analysis_year <- 2024L
candidate_pool_input <- uc_salary_path("data", "processed", paste0("uc_salary_", analysis_year, "_candidate_pool.csv"))
analytic_output <- uc_salary_path("data", "processed", paste0("uc_salary_", analysis_year, "_analytic_dataset.csv"))
cohort_summary_output <- uc_salary_path("data", "processed", paste0("uc_salary_", analysis_year, "_summary_by_cohort.csv"))
cohort_reference_output <- uc_salary_path("data", "processed", paste0("uc_salary_", analysis_year, "_cohort_reference.csv"))
title_crosswalk_output <- uc_salary_path("data", "processed", paste0("uc_salary_", analysis_year, "_title_crosswalk.csv"))
join_audit_output <- uc_salary_path("outputs", "logs", "joins", paste0("uc_salary_", analysis_year, "_title_join_audit.csv"))
qa_output <- uc_salary_path("outputs", "logs", "qa", paste0("uc_salary_", analysis_year, "_cleaning_qa_summary.csv"))

uc_salary_ensure_directory(dirname(join_audit_output))
```

<!-- 9. Validation -->

## Input validation for the classification audit

These checks confirm that the candidate pool and the rule/reference files needed to resolve overlap are present before any cohort assignment begins.

```{r validate-classification-inputs}
stopifnot(file.exists(candidate_pool_input))

candidate_pool <- readr::read_csv(candidate_pool_input, show_col_types = FALSE)
query_registry <- uc_salary_read_query_registry() %>%
  dplyr::filter(analysis_year == !!analysis_year)
title_reference <- uc_salary_read_title_reference()

stopifnot(nrow(candidate_pool) > 0)
stopifnot(nrow(query_registry) > 0)
stopifnot(nrow(title_reference) > 0)

validation_summary <- tibble::tibble(
  validation_check = c(
    "Candidate pool rows available",
    "2024 cohort rules available",
    "Academic title reference rows available"
  ),
  validation_result = c(
    nrow(candidate_pool),
    nrow(query_registry),
    nrow(title_reference)
  )
)

knitr::kable(validation_summary)
```

<!-- 10. Importing data -->

## Candidate-pool overlap entering classification

This section imports the candidate pool and reference data used to classify records so the overlap built into the upstream scrape remains visible before any deduplication or cohort assignment happens. The counts below are still upstream candidate-pool counts, not final cohort sizes.

### Candidate-pool overlap before classification

The table below shows the size of each upstream query return before priority rules and title filters resolve overlap. This is the baseline against which the classification audit should be read.

```{r inspect-upstream-candidate-pool}
candidate_pool %>%
  dplyr::count(query_slug, sort = TRUE) %>%
  knitr::kable()
```

<!-- 11. Processing data -->

## Construction of classified cohorts and audit outputs

This section applies the cohort rules, joins the academic title reference where available, resolves overlaps using explicit query priorities, and prepares the machine-readable processed outputs that document the cohort recodes. Analytically, it moves through four steps: applying the cohort rules, joining the academic title reference where available, resolving overlap with explicit priorities, and writing the audit tables that explain how the final cohort file was built.

```{r assign-cohorts-and-write-audits}
analytic_data <- candidate_pool %>%
  uc_salary_assign_cohorts(query_registry = query_registry) %>%
  uc_salary_join_title_reference(title_reference = title_reference) %>%
  dplyr::arrange(cohort_group, analysis_cohort, dplyr::desc(total_pay_benefits))

cohort_summary <- analytic_data %>%
  dplyr::group_by(cohort_group, analysis_cohort) %>%
  dplyr::summarise(
    workers = dplyr::n(),
    median_regular_pay = stats::median(regular_pay, na.rm = TRUE),
    median_total_pay = stats::median(total_pay, na.rm = TRUE),
    median_total_pay_benefits = stats::median(total_pay_benefits, na.rm = TRUE),
    mean_total_pay_benefits = mean(total_pay_benefits, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(cohort_group, dplyr::desc(median_total_pay_benefits))

title_join_audit <- analytic_data %>%
  dplyr::count(analysis_cohort, job_title_clean, title_reference_match, sort = TRUE)

cohort_reference <- query_registry %>%
  dplyr::select(
    analysis_year,
    priority,
    query_slug,
    search_term,
    analysis_cohort,
    cohort_group,
    cohort_family,
    include_regex,
    exclude_regex,
    rationale
  )

title_crosswalk <- analytic_data %>%
  dplyr::group_by(
    source_search_term,
    query_slug,
    job_title,
    job_title_clean,
    analysis_cohort,
    cohort_group,
    cohort_family,
    cohort_rule_rationale,
    uc_title_job_code,
    uc_cto_name,
    title_reference_match,
    police_role_group
  ) %>%
  dplyr::summarise(
    source_records = dplyr::n(),
    .groups = "drop"
  ) %>%
  dplyr::arrange(cohort_group, analysis_cohort, dplyr::desc(source_records), job_title_clean)

readr::write_csv(analytic_data, analytic_output)
readr::write_csv(cohort_summary, cohort_summary_output)
readr::write_csv(cohort_reference, cohort_reference_output)
readr::write_csv(title_crosswalk, title_crosswalk_output)
readr::write_csv(title_join_audit, join_audit_output)

cohort_summary %>%
  knitr::kable()
```

<!-- 12. Inspect data -->

## Classified cohorts after priority resolution

This section inspects the cleaned analytic dataset so cohort sizes, title patterns, and the post-priority cohort allocation are visible before the final analysis report uses the file. At this point the rows represent the repository's locked 2024 person-level cohorts rather than the broader overlapping scrape corpus.

```{r inspect-classified-cohorts}
analytic_data %>%
  dplyr::count(cohort_group, analysis_cohort, sort = TRUE) %>%
  knitr::kable()
```

<!-- 13. QA -->

## Report-level validation checks

This section records key quality checks for duplicates, missing compensation fields, and reference joins so downstream users can verify that the classification stage resolved overlap rather than silently carrying it forward. The title-reference match rate is reported as an audit statistic, not as a completeness claim for all UC staff titles.

```{r qa-classification-audit}
post_classification_duplicate_count <- analytic_data %>%
  dplyr::count(employee_name_clean, job_title_clean, year, agency_clean) %>%
  dplyr::filter(n > 1) %>%
  nrow()

qa_summary <- tibble::tibble(
  qa_metric = c(
    "Analytic dataset rows",
    "Distinct analysis cohorts",
    "Post-classification duplicate keys",
    "Rows with missing total compensation",
    "Academic title reference match rate"
  ),
  qa_value = c(
    nrow(analytic_data),
    dplyr::n_distinct(analytic_data$analysis_cohort),
    post_classification_duplicate_count,
    sum(is.na(analytic_data$total_pay_benefits)),
    mean(analytic_data$title_reference_match %in% c(TRUE, "TRUE"), na.rm = TRUE)
  )
)

readr::write_csv(qa_summary, qa_output)

qa_summary %>%
  knitr::kable()
```

<!-- 14. Export clean data -->

## Saved datasets, crosswalks, and audit artifacts

This artifact registry functions as a reproducibility appendix for the classification stage. It lists the cleaned analysis-ready artifacts that the final report consumes, including the cohort reference, title crosswalk, join audit, and QA outputs that document the cleaning logic explicitly.

```{r list-classification-artifacts}
export_paths <- tibble::tibble(
  artifact = c(
    "Analysis-ready 2024 cohort dataset",
    "2024 summary by cohort",
    "2024 cohort-rule reference table",
    "2024 title crosswalk (observed titles to analytic cohorts)",
    "2024 title-join audit (academic reference join)",
    "2024 classification QA summary"
  ),
  output_path = c(
    analytic_output,
    cohort_summary_output,
    cohort_reference_output,
    title_crosswalk_output,
    join_audit_output,
    qa_output
  )
)

export_paths %>%
  dplyr::mutate(output_path = uc_salary_relpath(output_path)) %>%
  knitr::kable()
```
