---
title: "UC Police and Librarian Payroll Comparison Over Time"
author: "Robert J. Dellinger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
bibliography: "../references/uc_salary_sources.bib"
link-citations: true
output:
  html_document:
    theme: flatly
    toc: true
    df_print: paged
---

# Title

This section states the title of the standalone police-versus-librarian payroll workflow.

UC Police and Librarian Payroll Comparison Over Time

# Name

This section identifies the analyst responsible for this workflow.

Robert J. Dellinger

# Date

This section records the run date for the workflow.

`r format(Sys.Date(), "%B %d, %Y")`

# Objective

This section explains the purpose of the workflow before any code runs.

The objective is to build a publication-quality, reusable University of California over-time comparison between police payroll and broad librarian-series payroll using reconstructed annual payroll estimates from the public Transparent California job-title summary pages and the repository's documented librarian title crosswalk.

# Source

This section names the source materials used in the workflow.

The workflow uses Transparent California's annual University of California job-title summary pages for `2011-2024` [@transparentcalifornia_uc_2026; @transparentcalifornia_uc_job_title_summary_2026], the latest-available 2024 University of California agency page that is still marked in progress [@transparentcalifornia_uc_2024_progress_2026], the exact police person-level series produced elsewhere in this repository for validation [@transparentcalifornia_uc_2026], and the repository's cleaned 2024 librarian title crosswalk to audit current librarian title strings against the title matches already validated in the person-level workflow.

# Source documentation and information

This section summarizes the source fields, cohort-definition rules, and interpretation limits that matter for the workflow.

Transparent California's job-title summary pages report `Count`, `Avg Regular pay`, `Avg Overtime pay`, `Avg Other pay`, `Avg Total pay`, `Avg Benefits`, and `Avg Total pay & benefits` for each title-year combination [@transparentcalifornia_uc_job_title_summary_2026]. Those pages also note that the averages are designed to capture full-year employees and exclude part-time, seasonal, and partial-year workers [@transparentcalifornia_uc_job_title_summary_2026]. Because of that note, this script treats title-summary payroll totals as reconstructed annual payroll estimates rather than exact audited payroll sums. The librarian side of the comparison is defined as a broad librarian-series cohort built from observed title-summary strings that contain `LIBRARIAN`, including university, law, associate, assistant, career, potential-career, non-represented, temporary, and visiting librarian titles. The repository's 2024 librarian crosswalk is used only as an audit aid for current title matching, while the actual over-time inclusion rule is the explicit title crosswalk written by this script. The resulting over-time figures are nominal dollars, not inflation-adjusted dollars, and the newest year should still be read as the latest available 2024 snapshot rather than a finalized University of California census [@transparentcalifornia_uc_2024_progress_2026].

# Methods, including equations and conceptual definitions

This section explains the core estimation and validation logic before the code executes.

The workflow treats each year as a separate repeated cross-section rather than averaging workers across years. For each title `t` in year `y`, the title-summary pages provide a worker count and an average compensation value. The workflow therefore constructs estimated annual payroll by title and then sums those title-level estimates within each cohort:

- `estimated_title_total_pay_(t,y) = count_(t,y) * avg_total_pay_(t,y)`
- `estimated_group_total_pay_y = Σ estimated_title_total_pay_(t,y)` for all titles `t` in the cohort
- `police_share_of_librarian_payroll_y = estimated_police_total_pay_y / estimated_librarian_total_pay_y`
- `librarian_payroll_multiple_of_police_y = estimated_librarian_total_pay_y / estimated_police_total_pay_y`
- `average_librarian_payroll_equivalents_y = estimated_police_total_pay_y / weighted_mean_librarian_total_pay_y`

Because the title-summary pages are not person-level payroll extracts, the workflow validates the police title-summary estimates against the exact police person-level totals from the repository's police series. Those validation gaps are reported directly and the librarian totals remain labeled as reconstructed annual payroll estimates rather than exact payroll totals.

# Setup and load packages

This section loads helper functions, packages, and project paths.

```{r setup-and-load-packages}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

utility_path <- if (file.exists("scripts/functions/uc_salary_utils.R")) {
  "scripts/functions/uc_salary_utils.R"
} else {
  "functions/uc_salary_utils.R"
}

source(utility_path)
uc_salary_load_packages()

analysis_start_year <- 2011L
year_status <- uc_salary_get_agency_year_status(agency_slug = "university-of-california")
analysis_end_year <- year_status$current_page_year[[1]]
analysis_years <- seq.int(analysis_start_year, analysis_end_year)
series_stub <- paste0("uc_police_vs_librarian_", analysis_start_year, "_", analysis_end_year)

police_detail_input <- uc_salary_path(
  "data",
  "processed",
  paste0("uc_police_spending_", analysis_start_year, "_", analysis_end_year, "_detail.csv")
)
librarian_crosswalk_input <- uc_salary_path("data", "processed", "uc_salary_2024_title_crosswalk.csv")

title_summary_detail_output <- uc_salary_path("data", "processed", paste0(series_stub, "_title_summary_detail.csv"))
title_group_output <- uc_salary_path("outputs", "tables", paste0(series_stub, "_title_group_reference.csv"))
annual_output <- uc_salary_path("outputs", "tables", paste0(series_stub, "_annual_summary.csv"))
comparison_output <- uc_salary_path("outputs", "tables", paste0(series_stub, "_comparison_summary.csv"))
validation_output <- uc_salary_path("outputs", "tables", paste0(series_stub, "_police_validation.csv"))
scrape_manifest_output <- uc_salary_path("outputs", "logs", "qa", paste0(series_stub, "_scrape_manifest.csv"))
qa_output <- uc_salary_path("outputs", "logs", "qa", paste0(series_stub, "_qa_summary.csv"))
figure_manifest_output <- uc_salary_path("outputs", "logs", "qa", paste0(series_stub, "_figure_manifest.csv"))

uc_salary_ensure_directory(dirname(title_summary_detail_output))
uc_salary_ensure_directory(dirname(title_group_output))
uc_salary_ensure_directory(dirname(scrape_manifest_output))
```

# Validation

This section validates that the required inputs and the latest-available time window are in place before scraping begins.

```{r validation}
stopifnot(!is.na(analysis_end_year))
stopifnot(analysis_end_year >= analysis_start_year)
stopifnot(length(analysis_years) > 0)
stopifnot(file.exists(police_detail_input))
stopifnot(file.exists(librarian_crosswalk_input))

librarian_crosswalk <- readr::read_csv(librarian_crosswalk_input, show_col_types = FALSE)
current_librarian_title_reference <- librarian_crosswalk %>%
  dplyr::filter(analysis_cohort == "Librarians") %>%
  dplyr::mutate(
    current_2024_title_job_code = dplyr::na_if(uc_title_job_code, "NA"),
    current_2024_cto_name = dplyr::na_if(uc_cto_name, "NA"),
    current_2024_title_reference_match = title_reference_match %in% c(TRUE, "TRUE")
  ) %>%
  dplyr::select(
    job_title_clean,
    current_2024_title_job_code,
    current_2024_cto_name,
    current_2024_title_reference_match
  ) %>%
  dplyr::distinct(job_title_clean, .keep_all = TRUE)

stopifnot(nrow(current_librarian_title_reference) > 0)

validation_summary <- tibble::tibble(
  validation_check = c(
    "Current Transparent California UC page year",
    "Current page marked as in progress",
    "Latest available year used in this script",
    "First year used in this script",
    "Number of years in the comparison series",
    "Exact police detail file available",
    "Current librarian crosswalk file available",
    "Current librarian crosswalk rows available"
  ),
  validation_result = c(
    year_status$current_page_year[[1]],
    year_status$data_collection_in_progress[[1]],
    analysis_end_year,
    analysis_start_year,
    length(analysis_years),
    file.exists(police_detail_input),
    file.exists(librarian_crosswalk_input),
    nrow(current_librarian_title_reference)
  )
)

knitr::kable(validation_summary)
```

# Importing data

This section documents the planned yearly title-summary inputs and the broad librarian-series rule that defines the comparison cohort.

```{r importing-data}
planned_input_manifest <- tibble::tibble(
  year = analysis_years,
  raw_file = file.path(
    uc_salary_path("data", "raw", "transparent_california"),
    as.character(analysis_years),
    paste0("tc_", analysis_years, "_job_title_summary.csv")
  )
) %>%
  dplyr::mutate(raw_file_exists = file.exists(raw_file))

librarian_series_rules <- tibble::tibble(
  analysis_cohort = "Librarians",
  include_regex = "LIBRARIAN",
  exclude_regex = NA_character_,
  rule_note = paste(
    "The librarian comparison uses a broad librarian-series payroll cohort built from every observed UC title-summary",
    "string that contains `LIBRARIAN`, including university, law, associate, assistant, career, potential-career,",
    "non-represented, temporary, and visiting librarian titles."
  )
)

planned_input_manifest %>%
  knitr::kable()

librarian_series_rules %>%
  knitr::kable()

current_librarian_title_reference %>%
  dplyr::count(current_2024_cto_name, current_2024_title_reference_match, sort = TRUE) %>%
  knitr::kable()
```

# Processing data

This section scrapes or reuses yearly title-summary pages, defines the police and librarian cohorts, validates the title-summary police estimates against the exact police series, and writes publication-ready figures and tables.

```{r processing-data}
yearly_title_summary_results <- purrr::map(
  analysis_years,
  function(year_value) {
    raw_year_directory <- uc_salary_path("data", "raw", "transparent_california", as.character(year_value))
    raw_output_path <- file.path(raw_year_directory, paste0("tc_", year_value, "_job_title_summary.csv"))

    uc_salary_ensure_directory(raw_year_directory)

    # Reuse cached yearly title-summary extracts when they already exist.
    year_title_summary <- if (file.exists(raw_output_path)) {
      cached_year_summary <- readr::read_csv(raw_output_path, show_col_types = FALSE)

      if (uc_salary_cache_is_current(cached_year_summary)) {
        cached_year_summary
      } else {
        raw_year_summary <- uc_salary_scrape_job_title_summary(
          year = year_value,
          pause_seconds = 0.10,
          verbose = TRUE
        )

        readr::write_csv(raw_year_summary, raw_output_path)
        raw_year_summary
      }
    } else {
      raw_year_summary <- uc_salary_scrape_job_title_summary(
        year = year_value,
        pause_seconds = 0.10,
        verbose = TRUE
      )

      readr::write_csv(raw_year_summary, raw_output_path)
      raw_year_summary
    }

    manifest_row <- tibble::tibble(
      year = year_value,
      raw_file = raw_output_path,
      title_rows_scraped = nrow(year_title_summary),
      unique_titles_scraped = dplyr::n_distinct(year_title_summary$job_title_clean),
      total_pages_reported = max(year_title_summary$source_total_pages, na.rm = TRUE),
      unique_pages_scraped = dplyr::n_distinct(year_title_summary$source_page),
      source_note = dplyr::first(year_title_summary$source_note)
    )

    list(data = year_title_summary, manifest = manifest_row)
  }
)

title_summary_detail <- dplyr::bind_rows(purrr::map(yearly_title_summary_results, "data"))
scrape_manifest <- dplyr::bind_rows(purrr::map(yearly_title_summary_results, "manifest"))

librarian_title_candidates <- title_summary_detail %>%
  dplyr::filter(stringr::str_detect(job_title_clean, "LIBRARIAN")) %>%
  dplyr::group_by(job_title, job_title_clean) %>%
  dplyr::summarise(
    first_year = min(year, na.rm = TRUE),
    last_year = max(year, na.rm = TRUE),
    max_worker_count = max(count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    analysis_cohort = "Librarians",
    included_in_comparison = TRUE,
    comparison_note = dplyr::case_when(
      stringr::str_detect(job_title_clean, "UNIV LIBRARIAN|UNIVERSITY LIBRARIAN") ~ "Included as part of the broad librarian series; this title captures university-librarian leadership positions within library payroll.",
      stringr::str_detect(job_title_clean, "LAW LIBRARIAN") ~ "Included as part of the broad librarian series; law-library titles remain within the public-facing librarian payroll comparison.",
      stringr::str_detect(job_title_clean, "TEMP|VIS") ~ "Included as part of the broad librarian series because the title-summary pages already aim to capture full-year workers, but this title should not be read as a permanent appointment count.",
      TRUE ~ "Included as part of the broad librarian series in the police-versus-librarian comparison."
    )
  ) %>%
  dplyr::left_join(
    current_librarian_title_reference,
    by = "job_title_clean"
  )

included_librarian_titles <- librarian_title_candidates %>%
  dplyr::filter(included_in_comparison, !is.na(analysis_cohort))

comparison_title_detail <- title_summary_detail %>%
  dplyr::left_join(
    included_librarian_titles %>%
      dplyr::select(
        job_title_clean,
        analysis_cohort,
        current_2024_title_job_code,
        current_2024_cto_name,
        current_2024_title_reference_match
      ),
    by = "job_title_clean"
  ) %>%
  dplyr::mutate(
    analysis_cohort = dplyr::case_when(
      stringr::str_detect(job_title_clean, "\\bPOLICE\\b") ~ "Police",
      TRUE ~ analysis_cohort
    ),
    cohort_group = dplyr::case_when(
      analysis_cohort == "Police" ~ "Police",
      analysis_cohort == "Librarians" ~ "Librarians",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::filter(!is.na(analysis_cohort))

annual_summary <- comparison_title_detail %>%
  dplyr::group_by(year, cohort_group, analysis_cohort) %>%
  dplyr::summarise(
    title_rows = dplyr::n(),
    workers = sum(count, na.rm = TRUE),
    estimated_regular_pay = sum(estimated_regular_pay, na.rm = TRUE),
    estimated_overtime_pay = sum(estimated_overtime_pay, na.rm = TRUE),
    estimated_other_pay = sum(estimated_other_pay, na.rm = TRUE),
    estimated_total_pay = sum(estimated_total_pay, na.rm = TRUE),
    estimated_benefits = sum(estimated_benefits, na.rm = TRUE),
    estimated_total_pay_benefits = sum(estimated_total_pay_benefits, na.rm = TRUE),
    weighted_avg_regular_pay = stats::weighted.mean(avg_regular_pay, w = count, na.rm = TRUE),
    weighted_avg_total_pay = stats::weighted.mean(avg_total_pay, w = count, na.rm = TRUE),
    weighted_avg_total_pay_benefits = stats::weighted.mean(avg_total_pay_benefits, w = count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(year, cohort_group, analysis_cohort)

police_exact_annual <- readr::read_csv(police_detail_input, show_col_types = FALSE) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(
    exact_police_workers = dplyr::n(),
    exact_police_total_pay = sum(total_pay, na.rm = TRUE),
    exact_police_total_pay_benefits = sum(total_pay_benefits, na.rm = TRUE),
    .groups = "drop"
  )

police_validation <- annual_summary %>%
  dplyr::filter(analysis_cohort == "Police") %>%
  dplyr::transmute(
    year,
    estimated_police_workers = workers,
    estimated_police_total_pay = estimated_total_pay,
    estimated_police_total_pay_benefits = estimated_total_pay_benefits
  ) %>%
  dplyr::left_join(police_exact_annual, by = "year") %>%
  dplyr::mutate(
    worker_gap = estimated_police_workers - exact_police_workers,
    worker_gap_pct = dplyr::if_else(exact_police_workers > 0, estimated_police_workers / exact_police_workers - 1, NA_real_),
    total_pay_gap = estimated_police_total_pay - exact_police_total_pay,
    total_pay_gap_pct = dplyr::if_else(exact_police_total_pay > 0, estimated_police_total_pay / exact_police_total_pay - 1, NA_real_),
    total_pay_benefits_gap = estimated_police_total_pay_benefits - exact_police_total_pay_benefits,
    total_pay_benefits_gap_pct = dplyr::if_else(
      exact_police_total_pay_benefits > 0,
      estimated_police_total_pay_benefits / exact_police_total_pay_benefits - 1,
      NA_real_
    )
  )

comparison_summary <- annual_summary %>%
  dplyr::filter(analysis_cohort %in% c("Police", "Librarians")) %>%
  dplyr::select(year, analysis_cohort, workers, estimated_total_pay, estimated_total_pay_benefits, weighted_avg_total_pay) %>%
  tidyr::pivot_wider(
    names_from = analysis_cohort,
    values_from = c(workers, estimated_total_pay, estimated_total_pay_benefits, weighted_avg_total_pay),
    names_sep = "__"
  ) %>%
  dplyr::arrange(year) %>%
  dplyr::mutate(
    police_share_of_librarian_payroll = `estimated_total_pay__Police` / `estimated_total_pay__Librarians`,
    librarian_payroll_multiple_of_police = `estimated_total_pay__Librarians` / `estimated_total_pay__Police`,
    average_librarian_payroll_equivalents = `estimated_total_pay__Police` / `weighted_avg_total_pay__Librarians`,
    cumulative_police_total_pay = cumsum(`estimated_total_pay__Police`),
    cumulative_librarian_total_pay = cumsum(`estimated_total_pay__Librarians`)
  )

readr::write_csv(title_summary_detail, title_summary_detail_output)
readr::write_csv(librarian_title_candidates, title_group_output)
readr::write_csv(annual_summary, annual_output)
readr::write_csv(comparison_summary, comparison_output)
readr::write_csv(police_validation, validation_output)
readr::write_csv(scrape_manifest, scrape_manifest_output)

payroll_colour_map <- c(
  "Police" = "#3A506B",
  "Librarians" = "#8B1E3F"
)

share_colour_map <- c(
  "Police as share of librarian payroll" = "#8B1E3F"
)

equivalent_colour_map <- c(
  "Average librarian payrolls" = "#D1495B"
)

start_share_label <- scales::label_percent(accuracy = 1)(
  comparison_summary$police_share_of_librarian_payroll[comparison_summary$year == analysis_start_year][[1]]
)
end_share_label <- scales::label_percent(accuracy = 1)(
  comparison_summary$police_share_of_librarian_payroll[comparison_summary$year == analysis_end_year][[1]]
)
min_share_label <- scales::label_percent(accuracy = 1)(min(comparison_summary$police_share_of_librarian_payroll, na.rm = TRUE))
max_share_label <- scales::label_percent(accuracy = 1)(max(comparison_summary$police_share_of_librarian_payroll, na.rm = TRUE))

payroll_comparison_plot_data <- comparison_summary %>%
  dplyr::transmute(
    year,
    Police = `estimated_total_pay__Police`,
    Librarians = `estimated_total_pay__Librarians`
  ) %>%
  tidyr::pivot_longer(
    cols = c(Police, Librarians),
    names_to = "comparison_group",
    values_to = "estimated_total_pay"
  )

share_plot_data <- comparison_summary %>%
  dplyr::transmute(
    year,
    `Police as share of librarian payroll` = police_share_of_librarian_payroll
  ) %>%
  tidyr::pivot_longer(
    cols = c(`Police as share of librarian payroll`),
    names_to = "share_measure",
    values_to = "share_value"
  )

equivalent_plot_data <- comparison_summary %>%
  dplyr::transmute(
    year,
    `Average librarian payrolls` = average_librarian_payroll_equivalents
  ) %>%
  tidyr::pivot_longer(
    cols = c(`Average librarian payrolls`),
    names_to = "equivalent_measure",
    values_to = "equivalent_value"
  )

plot_police_vs_librarian_payroll <- ggplot2::ggplot(
  payroll_comparison_plot_data,
  ggplot2::aes(x = year, y = estimated_total_pay, colour = comparison_group)
) +
  ggplot2::geom_line(linewidth = 1.3) +
  ggplot2::geom_point(size = 2.6) +
  ggplot2::geom_text(
    data = payroll_comparison_plot_data %>% dplyr::filter(year == analysis_end_year),
    ggplot2::aes(
      label = paste0(comparison_group, ": ", scales::label_dollar(scale = 1e-6, suffix = "M", accuracy = 0.1)(estimated_total_pay))
    ),
    hjust = 0,
    nudge_x = 0.25,
    size = 3.2,
    show.legend = FALSE
  ) +
  ggplot2::scale_colour_manual(values = payroll_colour_map) +
  ggplot2::scale_x_continuous(breaks = analysis_years, expand = ggplot2::expansion(mult = c(0.01, 0.14))) +
  ggplot2::scale_y_continuous(
    labels = scales::label_dollar(scale_cut = scales::cut_short_scale()),
    expand = ggplot2::expansion(mult = c(0.02, 0.08))
  ) +
  ggplot2::labs(
    title = paste0(
      "Estimated librarian payroll stayed above estimated police payroll, ",
      analysis_start_year,
      "-",
      analysis_end_year
    ),
    subtitle = "Nominal annual totals use reconstructed Transparent California title-summary payroll estimates for broad police and librarian title-series cohorts.",
    x = NULL,
    y = "Estimated annual total pay",
    colour = NULL,
    caption = "Nominal annual payroll estimates reconstructed from Transparent California title counts × annual averages, not exact person-level payroll sums."
  ) +
  uc_salary_theme() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))

plot_police_payroll_share <- ggplot2::ggplot(
  share_plot_data,
  ggplot2::aes(x = year, y = share_value, colour = share_measure)
) +
  ggplot2::geom_line(linewidth = 1.3) +
  ggplot2::geom_point(size = 2.4) +
  ggplot2::geom_text(
    data = share_plot_data %>% dplyr::filter(year == analysis_end_year),
    ggplot2::aes(label = scales::label_percent(accuracy = 0.1)(share_value)),
    hjust = 0,
    nudge_x = 0.25,
    size = 3.2,
    show.legend = FALSE
  ) +
  ggplot2::scale_colour_manual(values = share_colour_map) +
  ggplot2::scale_x_continuous(breaks = analysis_years, expand = ggplot2::expansion(mult = c(0.01, 0.14))) +
  ggplot2::scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    expand = ggplot2::expansion(mult = c(0, 0.10))
  ) +
  ggplot2::labs(
    title = paste0(
      "Estimated police payroll stayed below estimated librarian payroll, ",
      analysis_start_year,
      "-",
      analysis_end_year
    ),
    subtitle = glue::glue("The police-to-librarian payroll ratio ranged from about {min_share_label} to {max_share_label} and ended at {end_share_label} in {analysis_end_year}."),
    x = NULL,
    y = "Police payroll share",
    colour = NULL,
    caption = "Shares use nominal annual title-summary payroll estimates; police totals are also cross-checked against the exact person-level police series."
  ) +
  uc_salary_theme() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))

plot_librarian_equivalent_payrolls <- ggplot2::ggplot(
  equivalent_plot_data,
  ggplot2::aes(x = year, y = equivalent_value, colour = equivalent_measure)
) +
  ggplot2::geom_line(linewidth = 1.3) +
  ggplot2::geom_point(size = 2.4) +
  ggplot2::geom_text(
    data = equivalent_plot_data %>% dplyr::filter(year == analysis_end_year),
    ggplot2::aes(label = scales::number(equivalent_value, accuracy = 1)),
    hjust = 0,
    nudge_x = 0.25,
    size = 3.2,
    show.legend = FALSE
  ) +
  ggplot2::scale_colour_manual(values = equivalent_colour_map) +
  ggplot2::scale_x_continuous(breaks = analysis_years, expand = ggplot2::expansion(mult = c(0.01, 0.14))) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.10))) +
  ggplot2::labs(
    title = paste0(
      "Estimated annual police payroll equaled hundreds of\naverage librarian payrolls, ",
      analysis_start_year,
      "-",
      analysis_end_year
    ),
    subtitle = "Equivalent payroll counts divide annual police estimated total pay by same-year weighted average librarian title-summary total pay.",
    x = NULL,
    y = "Equivalent same-year average payrolls",
    colour = NULL,
    caption = "Equivalent counts use nominal annual title-summary payroll estimates and do not represent authorized librarian FTEs or hiring plans."
  ) +
  uc_salary_theme() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))

figure_manifest <- tibble::tibble(
  figure = c(
    "Estimated annual police versus librarian payroll",
    "Police payroll share of librarian payroll",
    "Librarian payroll-equivalent counts for police payroll"
  ),
  output_path = c(
    uc_salary_save_plot(plot_police_vs_librarian_payroll, paste0(series_stub, "_annual_estimated_payroll.png")),
    uc_salary_save_plot(plot_police_payroll_share, paste0(series_stub, "_payroll_share.png")),
    uc_salary_save_plot(plot_librarian_equivalent_payrolls, paste0(series_stub, "_librarian_equivalent_payrolls.png"))
  )
)

readr::write_csv(figure_manifest, figure_manifest_output)
```

# Inspect data

This section reviews the finished annual comparison tables and the police-validation results before the report concludes.

```{r inspect-data}
annual_summary %>%
  dplyr::filter(analysis_cohort %in% c("Police", "Librarians")) %>%
  dplyr::arrange(year, analysis_cohort) %>%
  knitr::kable()

police_validation %>%
  dplyr::select(year, estimated_police_total_pay, exact_police_total_pay, total_pay_gap_pct) %>%
  knitr::kable()
```

# QA

This section records the main quality checks for title coverage, validation gaps, and figure creation.

```{r qa}
qa_summary <- tibble::tibble(
  qa_metric = c(
    "Years analyzed",
    "Title-summary rows in comparison detail",
    "Police title rows included",
    "Included librarian observed titles",
    "Included librarian titles with current 2024 crosswalk match",
    "Maximum absolute police total-pay validation gap",
    "Mean absolute police total-pay validation gap",
    "Figures written"
  ),
  qa_value = c(
    length(analysis_years),
    nrow(comparison_title_detail),
    dplyr::n_distinct(comparison_title_detail$job_title_clean[comparison_title_detail$analysis_cohort == "Police"]),
    nrow(included_librarian_titles),
    sum(included_librarian_titles$current_2024_title_reference_match, na.rm = TRUE),
    max(abs(police_validation$total_pay_gap_pct), na.rm = TRUE),
    mean(abs(police_validation$total_pay_gap_pct), na.rm = TRUE),
    nrow(figure_manifest)
  )
)

readr::write_csv(qa_summary, qa_output)

qa_summary %>%
  knitr::kable()
```

# Export clean data

This section lists the tables, logs, and figures written by the workflow.

```{r export-clean-data}
export_manifest <- dplyr::bind_rows(
  tibble::tibble(
    artifact = c(
      "Processed title-summary detail",
      "Librarian title-group reference",
      "Annual comparison summary",
      "Police-versus-librarian comparison summary",
      "Police validation table",
      "Scrape manifest",
      "Figure manifest",
      "QA summary"
    ),
    output_path = c(
      title_summary_detail_output,
      title_group_output,
      annual_output,
      comparison_output,
      validation_output,
      scrape_manifest_output,
      figure_manifest_output,
      qa_output
    )
  ),
  figure_manifest %>%
    dplyr::rename(artifact = figure)
)

knitr::kable(export_manifest)
```

# Works Cited

This section is reserved for the bibliography generated from the report citations.
